$Z_3$ topological order in the quantum dimer-pentamer model

### 2017-07-27 qdpm_paper.tex
Added new version of Fig. 4, moved figures around, added citation to PRB

We may understand the emergence of the $Z_2$ topologically ordered phase of IGTs from the QDM and QFPLM from the following picture. The additional vertex configurations that are allowed in the IGTs relative to the models with a dimer number constraint  have U(1) charge $Q_v = \pm 2$ (see \tabref{tab:defects}). It is known that when a $U(1)$ gauge field is coupled to a dynamic charge $q$ matter field, this results in a $Z_q$ gauge theory when the matter field condenses~\cite{Fradkin1979}; as in the IGT's we have added dynamic charge 2 defects to the underlying $U(1)$ symmetric model, and we may expect a $Z_2$ gauge theory to result. We note that pentamers are charge $\pm3$ objects in the QDM (see \tabref{tab:defects}); thus below we investigate whether introducing pentamers into the QDM leads to an effective $Z_3$ gauge theory (see \appref{sec:Z3GT} for a review of $Z_3$ gauge theory).

@@

The Hamiltonian of the QDPM is defined by \Eqref{eq:HGQDM} including both plaquette flips as well as local dimer resonances that create and destroy pairs of pentamers, and allow single pentamers to translate on the same sub-lattice; these local dynamics involving pentamers are illustrated in  \tabref{tab:pentamer_moves}. Note that pentamers give dynamics to ``staggered" regions with no flippable plaquettes; thus we expect the QDPM to have weaker correlations than the QDM, where the dynamics are more constrained. We may write $\HQDPM$ as the following:
\begin{equation}
\HQDPM \equiv \HQDM + H_{\rm{pen}}  \notag
\end{equation}
where $H_{\rm{pen}}$ only involves the pentamer resonances in \tabref{tab:pentamer_moves} and those related by symmetry:
\begin{align}
H_{\rm{pen}} \equiv \HGQDM \biggl( &\bigl\{ \left(c_{\nu_1},\bar{c}_{\nu_1}\right),\left(c_{\nu_2},\bar{c}_{\nu_2}\right),\left(c_{\nu_3},\bar{c}_{\nu_3}\right) \bigr\},\biggr.\notag\\
\biggl. &\bigl\{ \left(t_1,v_1\right),\left(t_2,v_2\right),\left(t_3,v_3\right) \bigr\} \biggr)  \label{eq:Hpen}
\end{align}

@@

@@

We explicitly demonstrate that the extensive topological degeneracy of the QDM is reduced to a finite degeneracy in the QDPM by computing the histogram of $U(1)$ topological winding numbers $\Phi_{X}$ and $\Phi_{Y}$, which is shown in \figref{fig:u1_wind_qdpm}. The different colors represent calculations that were initialized to different winding sectors $(\Phi_X,\Phi_Y)$, and we find that $\Phi_X$ and $\Phi_Y$ are not  strictly conserved (as they are in the QDM), but are only conserved modulo $3$. The histogram $P(\Phi)$ has a Gaussian shape with a width that increases with system size; this is expected as $\Phi = \pm L/2$ correspond to ordered configurations and thus there are fewer dimer configurations near these limits. We find no evidence of a lack of ergodicity within each of the 9 topological sectors, although we cannot rule out the possibility of isolated ``frozen" configurations which are not accessed by the dynamics described in \tabref{tab:pentamer_moves}; in particular we note that the QDPM gives dynamics to ``staggered" dimer configurations that are frozen in the QDM~\cite{Moessner2008}. For the remainder of the paper we present results in the $(0,0)$ topological sector, but we see no qualitative dependence between topological sectors for any of the results presented here.  Thus we find that the QDPM has a \emph{finite} $9$-fold degeneracy on the torus, where the topological sectors are labeled $(W_X^{Z_3},W_Y^{Z_3})$, in contrast with the extensive degeneracy of the QDM.
%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[]
    \centering
    \includegraphics[width=1.0\linewidth]{u1_wind_qdpm.pdf}
    \caption{The histogram of the $U(1)$ topological winding numbers $\Phi_X$ and $\Phi_Y$ as measured in the ground state of the RK point of the QDPM for three different lattice sizes. The colors correspond to calculations initialized to different topological sectors $(\Phi_X,\Phi_Y)$: blue $(0,0)$, red $(1,1)$, and purple $(-1,-1)$. The lines are fits to a Gaussian for each system size. Note that $\Phi$ is conserved modulo $3$, resulting in a $9$-fold topological degeneracy on a torus.}
    \label{fig:u1_wind_qdpm}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%

@@

We will now present a numerical study of the dimer density correlations in the RK ground state of the QDPM. In \figref{fig:dimer_3way} (a) we plot the bare dimer-dimer density correlation function $\langle n_{\bm{0}} n_{\bm{r}}\rangle$ on a $L=32$ lattice; here the origin $\bm{0}$ is represented by the orange link, and $\bm{r}$ varies over all other links of the lattice. These bare dimer correlations display only short range correlations with no obvious symmetry breaking. In \figref{fig:dimer_3way} (b) we show the equal-time dimer-dimer structure factor, 
\begin{equation}
S_d\left( \bm{k} \right) \equiv \frac{1}{ L^4} \sum_{i,j} e^{-i \bm{k} \cdot \left(\bm{r}_i - \bm{r}_j\right)} \left \langle n_{\bm{r}_i} n_{\bm{r}_j} \right \rangle  \notag
\end{equation}
where we restrict the sum to horizontal links, for a $L=64$ lattice. 
\figref{fig:dimer_3way} (c) shows $S_d(\bm{k})$ for a particular path through the Brillouin zone. The dimer structure factor displays broad maxima at $\bm{k} = (0, \pm \pi)$ and $\bm{k} = (\pm \pi, \pm \pi)$. The lack of rotational symmetry of $S_d(\bm{k})$ is due to the fact that we have isolated horizontal dimers; the correlations between vertical dimers display the same behavior, rotated by $\pi/2$.
We have plotted the finite size scaling of the intensity of the broad maximum of $S_d$ in  \figref{fig:bragg_scaling}, which we find to vanish in the thermodynamic limit; these maxima are therefore consistent with finite size effects due to short distance correlations. In \appref{sec:dimer_order_parameters} we additionally have plotted order parameters that specifically test for rotational and translational symmetry breaking; in all cases we find no symmetry breaking in the thermodynamic limit.

%%%%%%%%%%%%%%%%%%%%%%%
    \begin{figure}[t]
        \centering
        \includegraphics[width=1.0\columnwidth]{dimer_3way.pdf}
        \caption
        {
            (a) Gray scale of the dimer-dimer correlation function $\langle n_{\bm{0}}
            n_{\bm{r}}\rangle$ on an $L=32$ lattice. The origin corresponds to the orange link, and
            $\bm{r}$ varies over all other links of the lattice. Red links have correlations that are off
            the given scale. 
            (b) The horizontal dimer density structure factor $S_d(\bm{k})$ for the
            QDPM RK ground state on an $L=64$ lattice. The lack of $x\leftrightarrow y$ rotational
            symmetry is due to the fact that horizontal links have been isolated. 
            (c) The horizontal dimer-dimer structure factor $S_d(\bm{k})$ shown along a
            path in $k$ space for an $L=64$ lattice. Statistical errors are smaller than the symbols.
        }
        \label{fig:dimer_3way}
    \end{figure}
%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
    \centering
    \includegraphics[width=1.0\columnwidth]{peak_scaling.pdf}
    \caption{Finite-size scaling of the maxima in the dimer ($S_d$) and magnetic string ($S_M$) structure factors. The lines are fits to $L^{-2}$ scaling. All maxima vanish in the thermodynamic limit, indicating that the RK state of the QDPM is a symmetric dimer liquid. }
    \label{fig:bragg_scaling}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%

Additionally we have studied the nature of the decay of the connected dimer-dimer correlation function,
  \begin{equation}
    C_{\mathrm{d}} \left(r\right) \equiv \langle n_{\bm{0}} n_{\bm{r}} \rangle - \langle n_{\bm{0}} \rangle   \langle n_{\bm{r}} \rangle   \notag
\end{equation}
where $r$ is the distance between the links $\bm{0}$ and $\bm{r}$. In \figref{fig:spatial_dmr_cor}, $C_d$ between parallel links is plotted as a function of separation, for $\bm{r}$ perpendicular to the orientation of the dimers for a $L=64$ lattice. We find that $C_d$ decays exponentially with distance having a correlation length  $\chi\approx 1$. Additionally, $C_d$ displays a sub-lattice oscillation that is consistent with the finite-size maximum in $S_d$ at $(0,\pi)$. Thus we find that the RK point of the QDPM is a symmetric dimer liquid with exponentially decaying dimer correlations, in contrast with the power-law decaying correlations of the QDM.

%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
    \centering
    \includegraphics[width=1.0\columnwidth]{spatial_cors_parallel.pdf}
    \caption{The dimer-dimer density correlation function $C_d$ between parallel links in the QDPM RK ground state on a $L=64$ lattice, along the direction perpendicular to the link orientation.  The lines represent the best fit to an exponential decay with correlation length $\chi\approx 1$.}
    \label{fig:spatial_dmr_cor}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%

@@

To investigate the $Z_3$ gauge theory description of the QDPM RK ground state, we study the $Z_3$ magnetic string correlation function as defined in \eqref{eq:Sgamma} \& \eqref{eq:SQDPM}:
\begin{equation}
C_M \left( \bm{r}_{pp'} \right) \equiv \left \langle \exp \left[ \frac{2 \pi}{3} i \left( \Phi_{\gamma_{pp'}} -\tilde{\Phi}_{\gamma_{pp'}} \right) \right] \right \rangle  \notag
\end{equation}
where $p$ and $p'$ are two plaquettes, $\bm{r}_{pp'}$ is the displacement between $p$ and $p'$,  and $\gamma_{pp'}$ is any open string on the dual lattice
connecting $p$ and $p'$. In \figref{fig:vison_3way} (a), we have plotted $|C_M|$ for a $L=32$ lattice. Note that while $C_M$ depends on the choice of the background dimerization which fixes the gauge, $|C_M|$ is independent of the background dimerization. We find only short range magnetic string correlations that display no obvious symmetry breaking.

Additionally, we have computed the magnetic string structure factor
\begin{equation}
S_M\left( \bm{k} \right) \equiv \frac{1}{ L^4} \sum_{p,p'} e^{-i \bm{k} \cdot \bm{r}_{pp'}} C_M\left( \bm{r}_{pp'}\right)  \notag
\end{equation}
and plot the magnitude $|S_M|$ in \figref{fig:vison_3way} (b). We have chosen a columnar background dimerization comprising horizontal background dimers stacked in columns along the $y$-axis, as shown in \figref{fig:orientedSL}; this choice breaks the translational symmetry of the lattice and thus requires doubling the unit cell the $x$-direction. While $|C_M|$ displays no obvious symmetry breaking, there are peaks in $|S_M|$ at $(0,0)$ \& $(0,\pi$). We have plotted the finite-size scaling of these peaks in \figref{fig:bragg_scaling} which demonstrates that these peaks vanish in the thermodynamic limit. We have additionally studied $S_M$ using different background dimerization and find that the location of the finite-size peaks depend on the gauge choice; however in all cases we find that these peaks vanish in the thermodynamic limit. The vanishing of the $(0,\pi)$ peak demonstrates the restoration of rotational and translational symmetries in the thermodynamic limit, and the additional vanishing of the $(0,0)$ peak demonstrates the lack of magnetic vortex condensation, which is a mechanism for destroying topological order and breaking lattice symmetries~\cite{Jalabert1991,Ralko2007,Huh2011}. \figref{fig:vison_cor} shows the decay of $|C_M|$ along one of the axes of the dual lattice which we find decays exponentially with distance, with a correlation length $\chi\approx 3$. These exponentially decaying magnetic string correlations are therefore consistent with a symmetric dimer liquid with no magnetic vortex condensation.

%%%%%%%%%%%%%%%%%%%%%%%
    \begin{figure}
        \centering
        \includegraphics[width=1.0\columnwidth]{vison_3way.pdf}
        \caption
        {
            (a) Color scale of the $Z_3$ magnetic string correlation function $|C_M(\bm{r}_{\bm{0}p})|$,
            on a $L=32$ lattice where $\bm{0}$ corresponds to plaquette represented by the orange circle.
            (b) The magnitude of the $Z_3$ magnetic string structure factor $S_M$ for a $L=64$ lattice, using a columnar background dimerization as shown in \figref{fig:orientedSL}.             
            (c) $|S_M|$ shown along a path in $k$ space. Statistical errors are smaller than the symbols.
        }
        \label{fig:vison_3way}
    \end{figure}
%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
    \centering
    \includegraphics[width=1.0\columnwidth]{spatial_cors_z3_vis.pdf}
    \caption{ The magnitude of the $Z_3$ magnetic string correlation $|C_M|$ along one axis of the dual lattice of a $L=64$ lattice. The line corresponds to an exponential fit with correlations length $\chi\approx 3$.}
    \label{fig:vison_cor}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%

@@

To study the confinement of $Z_3$ electrically charged defects in the QDPM, we have Monte Carlo sampled the RK ground state doped with a single pair of monomers. We initialize the system with a monomer on each sub-lattice and supplement the local Monte Carlo updates with updates of the form:
\begin{equation}
\monomermvA  \Longleftrightarrow \monomermvB  \notag
\end{equation}
which appear to provide ergodic dynamics in the monomer-dimer-pentamer configuration space. Notice that monomers remain on the same sub-lattice, which is consistent with their assignment of a $Z_3$ electric charge of $\pm1$ (see \tabref{tab:defects}) and the positive and negative monomers live on different sub-lattices. We then can test for the confinement of monomers by studying the monomer-monomer density correlation function $\langle m_{\bm{0}} m_{\bm{r}} \rangle$, which is plotted in \figref{fig:monomer_correlations}. If the monomers are confined, we expect $\langle m_{\bm{0}} m_{\bm{r}} \rangle$ to decay to zero at large distances; conversely if monomers are deconfined, we expect the monomers to be randomly distributed on the lattice, such that $\langle m_{\bm{0}} m_{\bm{r}} \rangle \sim L^{-4}$ at large separations. \figref{fig:monomer_correlations} demonstrates that $\langle m_{\bm{0}} m_{\bm{r}} \rangle$ decays to $L^{-4}$ for each system size, indicating a deconfinement of monomers. This is in contrast with the QDM where monomer correlations decay as a power-law. Thus we find that $Z_3$ electric charges are deconfined in the QDPM.

%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%
    \begin{figure}[]
        \centering
        \includegraphics[width=1.0\columnwidth]{monomer_cor_log.pdf}
        \caption{ The monomer-monomer correlation function $\langle m_{\bm{0}} m_{\bm{r}} \rangle$ along a coordinate axis in the QDPM RK ground state doped with a single pair of monomers on lattices up to $L=64$. The horizontal lines correspond to $L^{-4}$ for the system size corresponding to the color.}
        \label{fig:monomer_correlations}
    \end{figure}
%%%%%%%%%%%%%%%%%%%%%%%

>>>>

We may understand the emergence of the $Z_2$ topologically ordered phase of IGTs from the QDM and QFPLM from the following picture. The additional vertex configurations that are allowed in the IGTs relative to the models with a dimer number constraint  have U(1) charge $Q_v = \pm 2$ (see \tabref{tab:defects}). It is known that when a $U(1)$ gauge field is coupled to a dynamic charge $q$ matter field, this results in a $Z_q$ gauge theory when the matter field condenses~\cite{Fradkin1979}; as in the IGT's we have added dynamic charge 2 defects to the underlying $U(1)$ symmetric model, and we may expect a $Z_2$ gauge theory to result. We note that pentamers are charge $\pm3$ objects in the QDM (see \tabref{tab:defects}); thus below we investigate whether introducing pentamers into the QDM leads to an effective $Z_3$ gauge theory, as suggested in Ref. \onlinecite{Herdman2013} (see \appref{sec:Z3GT} for a review of $Z_3$ gauge theory).

@@

The Hamiltonian of the QDPM is defined by \Eqref{eq:HGQDM} including both plaquette flips as well as local dimer resonances that create and destroy pairs of pentamers, and allow single pentamers to translate on the same sub-lattice; these local dynamics involving pentamers are illustrated in  \tabref{tab:pentamer_moves}. Note that pentamers give dynamics to ``staggered" regions with no flippable plaquettes that are frozen in the QDM (see \figref{fig:stag_to_flip} for an illustration of this); thus we expect the QDPM to have weaker correlations than the QDM, where the dynamics are more constrained. We may write $\HQDPM$ as the following:
\begin{equation}
\HQDPM \equiv \HQDM + H_{\rm{pen}}  \notag
\end{equation}
where $H_{\rm{pen}}$ only involves the pentamer resonances in \tabref{tab:pentamer_moves} and those related by symmetry:
\begin{align}
H_{\rm{pen}} \equiv \HGQDM \biggl( &\bigl\{ \left(c_{\nu_1},\bar{c}_{\nu_1}\right),\left(c_{\nu_2},\bar{c}_{\nu_2}\right),\left(c_{\nu_3},\bar{c}_{\nu_3}\right) \bigr\},\biggr.\notag\\
\biggl. &\bigl\{ \left(t_1,v_1\right),\left(t_2,v_2\right),\left(t_3,v_3\right) \bigr\} \biggr)  \label{eq:Hpen}
\end{align}

@@

 %%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[]
    \centering
    \includegraphics[width=1\linewidth]{StagToFlip.pdf}
    \caption{ An illustration of how the QDPM gives dynamics to dimer configurations that are frozen in the QDM. The transitions from panels (a)-(e) involve the pentamer moves $\nu_1$ and $\nu_2$ from \tabref{tab:pentamer_moves}; the transitions from (e)-(g) involve flipping dimers around the highlighted plaquettes. }
    \label{fig:stag_to_flip}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%

@@

We explicitly demonstrate that the extensive topological degeneracy of the QDM is reduced to a finite degeneracy in the QDPM by computing the histogram of $U(1)$ topological winding numbers $\Phi_{X}$ and $\Phi_{Y}$, which is shown in \figref{fig:u1_wind_qdpm}. The different colors represent calculations that were initialized to different winding sectors $(\Phi_X,\Phi_Y)$, and we find that $\Phi_X$ and $\Phi_Y$ are not  strictly conserved (as they are in the QDM), but are only conserved modulo $3$. The histogram $P(\Phi)$ has a Gaussian shape with a width that increases with system size; this is expected as $\Phi = \pm L/2$ correspond to ordered configurations and thus there are fewer dimer configurations near these limits. We find no evidence of a lack of ergodicity within each of the 9 topological sectors, although we cannot rule out the possibility of isolated ``frozen" configurations which are not accessed by the dynamics described in \tabref{tab:pentamer_moves}; in particular we note that the QDPM gives dynamics to ``staggered" dimer configurations that are frozen in the QDM (see \figref{fig:stag_to_flip}). For the remainder of the paper we present results in the $(0,0)$ topological sector, but we see no qualitative dependence between topological sectors for any of the results presented here.  Thus we find that the QDPM has a \emph{finite} $9$-fold degeneracy on the torus, where the topological sectors are labeled $(W_X^{Z_3},W_Y^{Z_3})$, in contrast with the extensive degeneracy of the QDM.

%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[]
    \centering
    \includegraphics[width=1.0\linewidth]{u1_wind_qdpm.pdf}
    \caption{The histogram of the $U(1)$ topological winding numbers $\Phi_X$ and $\Phi_Y$ as measured in the ground state of the RK point of the QDPM for three different lattice sizes. The colors correspond to calculations initialized to different topological sectors $(\Phi_X,\Phi_Y)$: blue $(0,0)$, red $(1,1)$, and purple $(-1,-1)$. The lines are fits to a Gaussian for each system size. Note that $\Phi$ is conserved modulo $3$, resulting in a $9$-fold topological degeneracy on a torus.}
    \label{fig:u1_wind_qdpm}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%

@@

We will now present a numerical study of the dimer density correlations in the RK ground state of the QDPM. In \figref{fig:dimer_3way} (a) we plot the bare dimer-dimer density correlation function $\langle n_{\bm{0}} n_{\bm{r}}\rangle$ on a $L=32$ lattice; here the origin $\bm{0}$ is represented by the orange link, and $\bm{r}$ varies over all other links of the lattice. These bare dimer correlations display only short range correlations with no obvious symmetry breaking. In \figref{fig:dimer_3way} (b) we show the equal-time dimer-dimer structure factor, 
\begin{equation}
S_d\left( \bm{k} \right) \equiv \frac{1}{ L^4} \sum_{i,j} e^{-i \bm{k} \cdot \left(\bm{r}_i - \bm{r}_j\right)} \left \langle n_{\bm{r}_i} n_{\bm{r}_j} \right \rangle  \notag
\end{equation}
where we restrict the sum to horizontal links, for a $L=64$ lattice. 
\figref{fig:dimer_3way} (c) shows $S_d(\bm{k})$ for a particular path through the Brillouin zone. The dimer structure factor displays broad maxima at $\bm{k} = (0, \pm \pi)$ and $\bm{k} = (\pm \pi, \pm \pi)$. The lack of rotational symmetry of $S_d(\bm{k})$ is due to the fact that we have isolated horizontal dimers; the correlations between vertical dimers display the same behavior, rotated by $\pi/2$.
We have plotted the finite size scaling of the intensity of the broad maximum of $S_d$ in  \figref{fig:bragg_scaling}, which we find to vanish in the thermodynamic limit; these maxima are therefore consistent with finite size effects due to short distance correlations. In \appref{sec:dimer_order_parameters} we additionally have plotted order parameters that specifically test for rotational and translational symmetry breaking; in all cases we find no symmetry breaking in the thermodynamic limit. 
%%%%%%%%%%%%%%%%%%%%%%%
    \begin{figure}[t]
        \centering
        \includegraphics[width=1.0\columnwidth]{dimer_3way.pdf}
        \caption
        {
            (a) Gray scale of the dimer-dimer correlation function $\langle n_{\bm{0}}
            n_{\bm{r}}\rangle$ on an $L=32$ lattice. The origin corresponds to the orange link, and
            $\bm{r}$ varies over all other links of the lattice. Red links have correlations that are off
            the given scale. 
            (b) The horizontal dimer density structure factor $S_d(\bm{k})$ for the
            QDPM RK ground state on an $L=64$ lattice. The lack of $x\leftrightarrow y$ rotational
            symmetry is due to the fact that horizontal links have been isolated. 
            (c) The horizontal dimer-dimer structure factor $S_d(\bm{k})$ shown along a
            path in $k$ space for an $L=64$ lattice. Statistical errors are smaller than the symbols.
        }
        \label{fig:dimer_3way}
    \end{figure}
%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[h!]
    \centering
    \includegraphics[width=1.0\columnwidth]{peak_scaling.pdf}
    \caption{Finite-size scaling of the maxima in the dimer ($S_d$) and magnetic string ($S_M$) structure factors. The lines are fits to $L^{-2}$ scaling. All maxima vanish in the thermodynamic limit, indicating that the RK state of the QDPM is a symmetric dimer liquid. }
    \label{fig:bragg_scaling}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%

Additionally we have studied the nature of the decay of the connected dimer-dimer correlation function,
  \begin{equation}
    C_{\mathrm{d}} \left(r\right) \equiv \langle n_{\bm{0}} n_{\bm{r}} \rangle - \langle n_{\bm{0}} \rangle   \langle n_{\bm{r}} \rangle   \notag
\end{equation}
where $r$ is the distance between the links $\bm{0}$ and $\bm{r}$. In \figref{fig:spatial_dmr_cor}, $C_d$ between parallel links is plotted as a function of separation, for $\bm{r}$ perpendicular to the orientation of the dimers for a $L=64$ lattice. We find that $C_d$ decays exponentially with distance having a correlation length  $\chi\approx 1$. Additionally, $C_d$ displays a sub-lattice oscillation that is consistent with the finite-size maximum in $S_d$ at $(0,\pi)$. Thus we find that the RK point of the QDPM is a symmetric dimer liquid with exponentially decaying dimer correlations, in contrast with the power-law decaying correlations of the QDM.
%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[h]
    \centering
    \includegraphics[width=1.0\columnwidth]{spatial_cors_parallel.pdf}
    \caption{The dimer-dimer density correlation function $C_d$ between parallel links in the QDPM RK ground state on a $L=64$ lattice, along the direction perpendicular to the link orientation.  The lines represent the best fit to an exponential decay with correlation length $\chi\approx 1$.}
    \label{fig:spatial_dmr_cor}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%

@@

To investigate the $Z_3$ gauge theory description of the QDPM RK ground state, we study the $Z_3$ magnetic string correlation function as defined in \eqref{eq:Sgamma} \& \eqref{eq:SQDPM}:
\begin{equation}
C_M \left( \bm{r}_{pp'} \right) \equiv \left \langle \exp \left[ \frac{2 \pi}{3} i \left( \Phi_{\gamma_{pp'}} -\tilde{\Phi}_{\gamma_{pp'}} \right) \right] \right \rangle  \notag
\end{equation}
where $p$ and $p'$ are two plaquettes, $\bm{r}_{pp'}$ is the displacement between $p$ and $p'$,  and $\gamma_{pp'}$ is any open string on the dual lattice
connecting $p$ and $p'$. In \figref{fig:vison_3way} (a), we have plotted $|C_M|$ for a $L=32$ lattice. Note that while $C_M$ depends on the choice of the background dimerization which fixes the gauge, $|C_M|$ is independent of the background dimerization. We find only short range magnetic string correlations that display no obvious symmetry breaking. 
%%%%%%%%%%%%%%%%%%%%%%%
    \begin{figure}
        \centering
        \includegraphics[width=1.0\columnwidth]{vison_3way.pdf}
        \caption
        {
            (a) Color scale of the $Z_3$ magnetic string correlation function $|C_M(\bm{r}_{\bm{0}p})|$,
            on a $L=32$ lattice where $\bm{0}$ corresponds to plaquette represented by the orange circle.
            (b) The magnitude of the $Z_3$ magnetic string structure factor $S_M$ for a $L=64$ lattice, using a columnar background dimerization as shown in \figref{fig:orientedSL}.             
            (c) $|S_M|$ shown along a path in $k$ space. Statistical errors are smaller than the symbols.
        }
        \label{fig:vison_3way}
    \end{figure}
%%%%%%%%%%%%%%%%%%%%%%%

Additionally, we have computed the magnetic string structure factor
\begin{equation}
S_M\left( \bm{k} \right) \equiv \frac{1}{ L^4} \sum_{p,p'} e^{-i \bm{k} \cdot \bm{r}_{pp'}} C_M\left( \bm{r}_{pp'}\right)  \notag
\end{equation}
and plot the magnitude $|S_M|$ in \figref{fig:vison_3way} (b). We have chosen a columnar background dimerization comprising horizontal background dimers stacked in columns along the $y$-axis, as shown in \figref{fig:orientedSL}; this choice breaks the translational symmetry of the lattice and thus requires doubling the unit cell the $x$-direction. While $|C_M|$ displays no obvious symmetry breaking, there are peaks in $|S_M|$ at $(0,0)$ \& $(0,\pi$). We have plotted the finite-size scaling of these peaks in \figref{fig:bragg_scaling} which demonstrates that these peaks vanish in the thermodynamic limit. We have additionally studied $S_M$ using different background dimerization and find that the location of the finite-size peaks depend on the gauge choice; however in all cases we find that these peaks vanish in the thermodynamic limit. The vanishing of the $(0,\pi)$ peak demonstrates the restoration of rotational and translational symmetries in the thermodynamic limit, and the additional vanishing of the $(0,0)$ peak demonstrates the lack of magnetic vortex condensation, which is a mechanism for destroying topological order and breaking lattice symmetries~\cite{Jalabert1991,Ralko2007,Huh2011}. \figref{fig:vison_cor} shows the decay of $|C_M|$ along one of the axes of the dual lattice which we find decays exponentially with distance, with a correlation length $\chi\approx 3$. These exponentially decaying magnetic string correlations are therefore consistent with a symmetric dimer liquid with no magnetic vortex condensation.
%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[]
    \centering
    \includegraphics[width=1.0\columnwidth]{spatial_cors_z3_vis.pdf}
    \caption{ The magnitude of the $Z_3$ magnetic string correlation $|C_M|$ along one axis of the dual lattice of a $L=64$ lattice. The line corresponds to an exponential fit with correlations length $\chi\approx 3$.}
    \label{fig:vison_cor}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%

@@

To study the confinement of $Z_3$ electrically charged defects in the QDPM, we have Monte Carlo sampled the RK ground state doped with a single pair of monomers. We initialize the system with a monomer on each sub-lattice and supplement the local Monte Carlo updates with updates of the form:
\begin{equation}
\monomermvA  \Longleftrightarrow \monomermvB  \notag
\end{equation}
which appear to provide ergodic dynamics in the monomer-dimer-pentamer configuration space. Notice that monomers remain on the same sub-lattice, which is consistent with their assignment of a $Z_3$ electric charge of $\pm1$ (see \tabref{tab:defects}) and the positive and negative monomers live on different sub-lattices. We then can test for the confinement of monomers by studying the monomer-monomer density correlation function $\langle m_{\bm{0}} m_{\bm{r}} \rangle$, which is plotted in \figref{fig:monomer_correlations}. If the monomers are confined, we expect $\langle m_{\bm{0}} m_{\bm{r}} \rangle$ to decay to zero at large distances; conversely if monomers are deconfined, we expect the monomers to be randomly distributed on the lattice, such that $\langle m_{\bm{0}} m_{\bm{r}} \rangle \simeq L^{-4}$ at large separations. \figref{fig:monomer_correlations} demonstrates that $\langle m_{\bm{0}} m_{\bm{r}} \rangle$ decays to $L^{-4}$ for each system size, indicating a deconfinement of monomers. This is in contrast with the QDM where monomer correlations decay as a power-law. Thus we find that $Z_3$ electric charges are deconfined in the QDPM.
%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%
    \begin{figure}[]
        \centering
        \includegraphics[width=1.0\columnwidth]{monomer_cor_log.pdf}
        \caption{ The monomer-monomer correlation function $\langle m_{\bm{0}} m_{\bm{r}} \rangle$ along a coordinate axis in the QDPM RK ground state doped with a single pair of monomers on lattices up to $L=64$. The horizontal lines correspond to $L^{-4}$ for the system size corresponding to the color.}
        \label{fig:monomer_correlations}
    \end{figure}
%%%%%%%%%%%%%%%%%%%%%%%
